\chapter*{Preface}

I would first like to thank Professor Dr. Marloes Maathuis for her
guidance in this project, as it provided invaluable clarity and direction
to my work. In addition to the time she set aside for our bi-weekly
meetings, her attendance at the group meetings organized
by her Ph.D. student Marco Eigenmann was greatly appreciated. These
meetings were critical in the development of more nuanced analysis and intuitive visualizations
through the discussions they facilitated between me and other students. I thank Marco Eigenmann for organizing them, and Jinzhou
Li, Armin Fingerle, Sanzio Monti, and Qikun Xiang for attending my presentations and listening attentively. A special thanks is
extended to C\'edric Bleutler and Leonard Henckel, both of whom were especially engaged and participated in lengthy discussions
both during and outside of these group meetings.

I would also like to acknowledge Professor Dr. Tilman Altwicker for his detailed literature suggestions for additional legal context
on peremptory challenges and Professor Dr. Samuel Baumgartner for his
research suggestions. The suggested literature provided an excellent
starting point for a first investigation of the topic. Of course, without the cooperation of Dr. Ronald Wright, Dr. George
Woodworth, Dr. Barbara O'Brien, and Dr. Catherine Grosso, all of whom generously provided me with data, the visualizations which
form the core of this work would not have been possible. As such, I am exceptionally grateful that they were so enthusiastic to
share the fruits of their labour to help cultivate mine, and hope that they find this work a fitting contribution to the
discussion of peremptory challenges.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "MasterThesisSfS"
%%% End: 
